---
title: "Deborah_bead_data"
author: "Biplabendu Das"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    code_folding: hide
    keep_md: no
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

set.seed(420)
rm(list = ls())

```


```{r housekeeping, include=FALSE}
## Load the libraries
pacman::p_load(tidyverse, viridis, conflicted)
pacman::p_load(ggplot2, patchwork)
# pacman::p_load(RSQLite, dbplyr)

## set conflict preference
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lmer", "lme4")
conflict_prefer("intersect", "dplyr")

## set path to your working directory
# path_to_repo = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data"
path_to_repo = "/Users/bidas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/02_bead_data"
path_to_figures = paste0(path_to_repo,"/results/figures/")

## load functions
# customized theme for publication quality figures
source(paste0(path_to_repo,"/functions/theme_publication.R"))
# # function to perform enrichment analysis
# source(paste0(path_to_repo,"/functions/enrichment_analysis.R"))
# # # function to plot z-scores (Cflo genes only)
# source(paste0(path_to_repo,"/functions/plot_zscores.R"))

bd.saveplot <- function(plot,name="myplot",width,height) {
  savingto <- paste0(path_to_figures,name,".png")
  writeLines(paste0("Plot is being saved to \n",savingto))
  png(savingto, 
    width = width, height = height, units = "cm", res = 300)
  print(plot)
  trash <- dev.off()
}

## checking pairwise dissimilarity using mean (median) and 95% CI
source(paste0(path_to_repo,"/functions/check_pairwise_significance.R"))
```


```{r set_parameters}
# minimum number of beads (out of 50) the colony must collect to be part of the analyses
atleast.beads <- 40

# color schemes
## year of obs
yr_colors <-c("grey60","grey60")
## age classes
age_colors <- c("#8D72A1", "#F26C7F", "#FBAD73")
## bead classes
# bead_colors <- c(slow="#1A80D9", fast="#F23030") # slow=blue, fast=red
bead_colors <- c(slow="#FBB369", fast="#3868A4")
## foraging classes
foraging_colors <- c("#9CD5A2", "#F29544")
```

# Overview/Goals

Deborah: 

This data set comes from an experiment with harvester ants last summer. Harvester ant foragers bring in lots of seeds and parts of plants, some of which are rejected by ants inside the nest, and then the refuse is carried out and put on the nest mound midden. 

> Here we presented small plastic beads coated in oil which the ants brought into the nest, and were later rejected and put back out on the midden.



# BEAD DATA

## Bead data 2021

### Part 1: INSPECT DATA

Two data sheets are available in the Excel workbook. For convenience of analysis, the excel sheets are saved individually as csv files and, if necessary, the column names for each were modified.

Sheet 1: foraging_data.csv
- contains three columns: date, colonyID, foraging_30s

Sheet 2: beads_out.csv
- contains XXX cols: num_out, date, colonyID, time_put_out, n_collected, n_returned_DayX (X %in% 1:6)

Brief description of the columns:
- date = date of observation
- colonyID = unique colony identifier
- time_put_out = time-of-day at which beads were put out
- n_collected = number of beads that had been brought into the nest at the end of foraging on Day 1
- n_returned_DayX = number of beads that had been brought *out of the nest* by the end of **Day X**

Note; 
NA means the colony is not active that day
NM means that the colony was doing nest maintenance, when they carry out refuse including beads

Some additional info:

Beads coated with oil were put out in 6 colonies on day 1, color green, then another 6 colonies on day 2, color green, then again with first 6 on day 3 color pink, then again with 2nd 6 on day 4 color pink, back to green days 5 and 6

Numbers of beads brought out are listed separately by color. E.g., colony 1131 took out 18 green beads on day 2 of green 8-24, and 3 pink beads on day 2 of pink 8-26.

For the 3rd round, for whatever number of green beads from the first round were not out by 8-27 or 8-28, then we don't know which day those were from.

### Part 2: Load data

```{r load_data}
# read foraging_data.csv
for.dat <- read.csv(
  paste0(path_to_repo,"/data/foraging_data.csv")
) %>% as_tibble() |> glimpse()

# read bead data
bead.dat <- read.csv(
  paste0(
    path_to_repo, 
    "/data/beads_out_new_24Jan25.csv"
  )
) %>% as_tibble() |> glimpse()

# read colony-age data
colony.dat <- read.csv(
# read.csv(
  paste0(
    path_to_repo, 
    "/data/foraging_counts_08_13_22.csv"
  )
) %>% 
  as_tibble() %>% 
  select(
    colonyID = colony_ID, 
    everything()
  ) %>% 
  mutate(
    colonyID = factor(colonyID),
    age_class = case_when(
      age_2022 <= 2 ~ "1-2_yr_old",
      age_2022 <= 4 ~ "3-4_yr_old",
      .default = "older_than_5_yr"
    ) |> 
      factor(
        levels = c(
          "1-2_yr_old", 
          "3-4_yr_old", 
          "older_than_5_yr"
        )
      )
  ) |> glimpse()
```


### Part 3: Summarize the data

```{r summarize_data, echo=F}

writeLines("What are the names of the colonies? \nShown as: colonyID")
bead.dat %>% 
  select(colonyID) %>%
  distinct() %>% 
  mutate(colonyID=as.factor(colonyID)) %>% 
  left_join(colony.dat, by="colonyID") %>%
  arrange(age_2022) %>% 
  head(4)


# Specify the dates for green/pink beads
green.days <- c(
  "8/23/21",
  "8/24/21",
  "8/27/21",
  "8/28/21"
) %>% 
  as.Date(format = "%m/%d/%y")
pink.days <- c(
  "8/25/21",
  "8/26/21"
) %>% 
  as.Date(format = "%m/%d/%y")
```

### Part 4: Clean Data

```{r clean_1}
df <-
  bead.dat %>%
  # arrange by colony IDs
  arrange(colonyID) %>% 
  ## FORMAT:
  mutate(
    date = as.Date(date, format = "%m/%d/%y"),
    colonyID = factor(colonyID),
    # define colony sets 
    colonyGroup = case_when(
      date %in% c(green.days[c(1,3)], pink.days[1]) ~ "batch_A",
      date %in% c(green.days[c(2,4)], pink.days[2]) ~ "batch_B",
      .default = NA
    ) |> 
      factor(
        levels = c("batch_A", "batch_B")
      ),
    # define replicates per colony set
    bead_color = case_when(
      date %in% green.days[1] ~ "rep1-green-21",
      date %in% green.days[2] ~ "rep2-green-21",
      date %in% pink.days[1]  ~ "rep3-pink-21",
      date %in% pink.days[2]  ~ "rep4-pink-21",
      date %in% green.days[3] ~ "rep5-green-21",
      date %in% green.days[4] ~ "rep6-green-21",
      .default = NA
    ) |> 
      as.factor()
  ) %>%
  ## SUBSET:
  # Keep data for first THREE days of observations
  select(
    num_out:n_returned_Day3, 
    bead_color, 
    colonyGroup
  ) %>%
  ## CLEAN:
  # format columns appropriately
  mutate(
    n_collected = readr::parse_number(n_collected),
    across(
      matches("^n_returned_Day"),
      ~ as.numeric(.x)
    )
  ) %>% 
  # keep colonies for which we have 
  filter(
    # data for at least one day
    if_any(
      matches("^n_returned_Day"),
      ~ !is.na(.x)
    ),
    # at least X number of collected beads 
    n_collected >= atleast.beads
  )

  
# query 1
# summarize the data
df %>% mutate_all(as.factor) |> summary() ## THERE is no NAs in the data

# query 1 (manuscript)
# summarize by colony set and bead_color
df %>% 
  group_by(colonyGroup, bead_color, date) %>%
  reframe(
    n_colonies = n()
  )

```

### Format 2021 data - I

```{r prep_data}
tidydat.2021 <-
  df %>% 
  rowwise() |> 
  # Cumulative prop 
  # create the cumulative sums
  mutate(
    cum_prop_day1 = n_returned_Day1/n_collected,
    cum_prop_day2 = sum(
      n_returned_Day1, 
      n_returned_Day2,
      na.rm = TRUE
    )/n_collected,
    cum_prop_day3 = sum(
      n_returned_Day1, 
      n_returned_Day2,
      n_returned_Day3,
      na.rm = TRUE
    )/n_collected
  ) %>% 
  ungroup() |> 
  tidyr::pivot_longer(
    cols = starts_with("cum_prop"),
    names_to = "day_of_obs",
    values_to = "cum_prop"
  ) %>% 
  mutate(
    day_of_obs  = readr::parse_number(day_of_obs),
    date        = date + day_of_obs - 1
  ) %>% 
  select(
    num_out, 
    date, 
    colonyID,
    n_collected, 
    bead_color, 
    colonyGroup,
    day_of_obs, 
    cum_prop
  )

```


___

### Format 2021 data - II

```{r abc}
old.summ <-
  tidydat.2021 %>% 
  left_join(
    colony.dat, 
    join_by(colonyID)
  ) %>% 
  mutate(
    age_obs = age_2022 - 1,
    day_of_obs = paste0(
      "D", day_of_obs, "_bN"
    ) |> 
      factor(
        levels = c("D1_bN", "D2_bN", "D3_bN")
      )
  ) %>% 
  ## change column names to make them compatible to the 2022 dataset
  select(
    colonyID, 
    date,
    bead_color,
    colonyGroup,
    age_obs, 
    age_class, 
    beads_collected = n_collected,
    counted_when = day_of_obs,
    cum_prop
  )

```

___


## Bead data 2022


```{r data_2022}
filename <- paste0(
  path_to_repo,
  "/data/bead_data_2022/all_bead_data_2022/2022_bead_data_master_replicate_one_two_three_29Aug22.csv")

## Read data
dat.2022 <- read.csv(filename,
                     header = T, stringsAsFactors = F,
                     na.strings = c(""," ","did_not_count","NA",
                                    "did_not_check",
                                    "it_rained_could_not_check")) %>% 
  as_tibble() %>% 
  
  # select the relevant columns
  select(date, colonyID, age_class, age_2022, 
         bead_color, beads_put_out, beads_collected,
         beads_returned_Day1_beforenoon,
         beads_returned_Day1_afternoon,
         beads_returned_Day2_morning,
         beads_returned_Day2_beforenoon,
         beads_returned_Day2_afternoon,
         beads_returned_Day3_beforenoon,
         beads_returned_Day4_morning,
         beads_returned_Day4_beforenoon) %>% 
  
  # calculate the cumulative proportion of beads returned
  group_by(date, bead_color) %>% 
  
  # filter out the colonies for which we still need to check the data
  filter(!(colonyID=="1560" & bead_color=="green")) %>% 
  filter(!(colonyID=="1288" & bead_color=="green")) %>%
  filter(!(colonyID=="1546" & bead_color=="green")) %>%
  
  # convert all the bead count columns as numeric
  mutate_at(vars(matches("beads_")), as.numeric) %>% 
  ungroup()
```


```{r prep_data_for_plot}
## Create a tidy dataset and calculate the cumulative proportion of beads returned by colonies
tidydat.2022a <-
  dat.2022 %>% 
  
  # replace all the NAs with 0 to be able to calculate proportions
  mutate_at(vars(matches("beads_")), function(x) ifelse(is.na(x), 0, x)) %>%
  
  # day 1 - before noon
  mutate(cum_prop_returned_D1_beforenoon = round(beads_returned_Day1_beforenoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D1_afternoon = round(cum_prop_returned_D1_beforenoon+beads_returned_Day1_afternoon/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_morning = round(cum_prop_returned_D1_afternoon+beads_returned_Day2_morning/beads_collected, 2)) %>% 
  mutate(cum_prop_returned_D2_beforenoon = round(cum_prop_returned_D2_morning+beads_returned_Day2_beforenoon/beads_collected,2)) %>%
  mutate(cum_prop_returned_D2_afternoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day2_afternoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D3_beforenoon = round(cum_prop_returned_D2_beforenoon+beads_returned_Day3_beforenoon/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_morning = round(cum_prop_returned_D3_beforenoon+beads_returned_Day4_morning/beads_collected,2)) %>% 
  mutate(cum_prop_returned_D4_beforenoon = round(cum_prop_returned_D4_morning+beads_returned_Day4_beforenoon/beads_collected,2)) %>% 
  
  arrange(age_2022) %>% 
  
  select(colonyID,
         age_2022, age_class,
         bead_color,
         cum_prop_returned_D1_beforenoon:cum_prop_returned_D4_beforenoon,
         everything())


tidydat.2022b <- 
  tidydat.2022a %>% 
  
  ## keep only colonies that collected at least 40 (of 50) beads offered
  filter(beads_collected >= atleast.beads) %>% 
  
  pivot_longer(cols = starts_with("cum_prop"),
               names_to = "counted_when",
               values_to = "cum_prop") %>% 
  
  ## format the ordinal columns correctly
  mutate(counted_when = ifelse(counted_when == "cum_prop_returned_D1_beforenoon", "D1_bN",
                            ifelse(counted_when == "cum_prop_returned_D1_afternoon", "D1_aN",
                                   ifelse(counted_when == "cum_prop_returned_D2_morning", "D2_M",
                                          ifelse(counted_when == "cum_prop_returned_D2_beforenoon", "D2_bN",
                                            ifelse(counted_when == "cum_prop_returned_D2_afternoon", "D2_aN",
                                            ifelse(counted_when == "cum_prop_returned_D3_beforenoon", "D3_bN",
                                            ifelse(counted_when == "cum_prop_returned_D4_morning", "D4_M",
                                                 "D4_bN")))))))) %>%
  mutate(counted_when = factor(counted_when, 
                               levels = c("D1_bN", "D1_aN", 
                                          "D2_M", "D2_bN", "D2_aN",
                                          "D3_bN",
                                          "D4_M", "D4_bN"))) %>% 
  mutate(bead_color = ifelse(bead_color %in% c("light_blue", "dark_blue"), "blue", bead_color))  %>% 
  mutate(colonyID = factor(colonyID), bead_color=factor(bead_color)) %>% 
  mutate(date = as.Date(date, format="%m/%d/%y")) %>% 
  
  mutate(age_obs = age_2022) %>% 
  
  # ## remove the timepoints that we do not have data for  
  # pink beads (round one)
  filter(!(bead_color=="pink" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_M")) %>%
  filter(!(bead_color=="pink" & counted_when == "D4_bN")) %>%
  # green beads (round two)
  filter(!(bead_color=="green" & counted_when == "D2_aN")) %>%
  filter(!(bead_color=="green" & counted_when == "D3_bN")) %>%
  filter(!(bead_color=="green" & counted_when == "D4_M")) %>%
  # blue beads (round three)
  filter(!(bead_color=="blue" & counted_when == "D1_aN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D2_bN")) %>%
  filter(!(bead_color=="blue" & counted_when == "D4_bN")) %>% 
  
  ## select columns to make them compatible with 2021 dataset
  select(colonyID, date,
         bead_color,
         age_2022, age_obs, age_class, 
         beads_collected, counted_when, cum_prop)
  
# Summarize the data
tidydat.2022b %>% summary()

# There are NAs in the bead_color column, check
tidydat.2022b %>% filter(is.na(bead_color))
# It appears that for colony 1504 and 1506, we do not have bead color infor for one of the replicates.
# Let's check to see if the other two color are already in the dataset, then by elimination the third and remaining color
# would be the correct entry.
tidydat.2022b %>% 
  filter(
    colonyID %in% c("1504","1506") & 
      counted_when %in% c("D1_bN","D2_M")
  ) %>% arrange(colonyID)
# Yes it is blue. Add the information
#
## CORRECTION HAPPENED HERE
#
tidydat.2022b[is.na(tidydat.2022b$bead_color),]$bead_color <- "blue"
```

### Create rep IDs

```{r}
# Rename the bead color values to reflect replicate
tidydat.2022 <- 
  tidydat.2022b %>% 
  mutate(
    bead_color = recode_factor(
      bead_color, 
      pink = "rep7-pink", 
      green = "rep8-green", 
      blue = "rep9-blue"
    ),
    colonyGroup = "batch_C"
  )
## Re-summarize the data
tidydat.2022 %>% summary()
```


### Filter 2022 dataset (Day 1-3)

This subsetted dataset contains data collected on Days 1-3, irrespective of
the time of day of collection:

M = early morning   ~ 0800 hrs
bN = before noon    ~ 1100 hrs
aN = after noon     ~ 1400 hrs 

```{r}
## Obtain data for Day 1-3
tidydat.2022.sub <-
  tidydat.2022 %>% 
    separate(
      counted_when, 
      c("day","counted_when"), 
      sep="_"
    ) %>% 
    filter(
      day %in% c("D1", "D2", "D3") #& 
        # counted_when %in% c("bN", "M")
    ) |> 
    group_by(colonyID, bead_color, day) %>% 
    reframe(
      cum_prop2=max(cum_prop),
      counted_when = paste(counted_when, collapse = ", "),
      n_counts = n()
    ) %>% 
    # mutate_all(as.factor) |> summary()
    mutate(
      counted_when = if_else(
        n_counts == 1,
        counted_when,
        case_when(
          stringr::str_detect(counted_when, "bN") ~ "bN",
          stringr::str_detect(counted_when, "aN") ~ "aN",
          .default = counted_when
        )
      )
    ) %>% 
    # janitor::tabyl(
    #   counted_when,
    #   counted_when2,
    #   day
    # )
    mutate(counted_when = paste0(day, "_", counted_when)) %>% 
    select(-day,-n_counts) %>%
    left_join(
      tidydat.2022, 
      by = c("colonyID","bead_color","counted_when")
    ) %>% 
    select(
      colonyID, 
      date, 
      bead_color,
      colonyGroup,
      age_obs, 
      age_class, 
      beads_collected, 
      counted_when, 
      cum_prop
    )

```


## Combine the 2021 and 2022 datasets

> Current dataset contains: All colonies from 2021 and 2022, for which we have counts of the number of beads returned by the colony on Day 1 and 2.

Note, for some colonies, day 2 data was observed in the morning (counted_when = D2_M) instead of later in the day (D2_bN).

```{r combine_datasets}
## Combine the two data sets
# 
tidydat.1 <-
  rbind(old.summ, tidydat.2022.sub) %>% 
  separate(
    counted_when, 
    c("day","counted_when"), 
    sep = "_"
  ) %>% 
  # glimpse() |> view()
  group_by(bead_color) %>% 
  mutate(
    rep = str_split(bead_color, "-")[[1]][1]
  ) %>%  
  ungroup()

## CLEANING STEP HAPPENING HERE
tidydat.1 %>%
  # filter(cum_prop>1) %>% 
  ## There are two values that are above 1, 
  ## and a closer inspection of those trials (see below)
  filter(
    colonyID %in% c("1440","1533") & 
      rep %in% c("rep8","rep9")
  )
  ## shows that in both cases, D1 value was very high and it is safe to assume a cum-prop of 1 on Day 2
tidydat <- tidydat.1 %>% 
  mutate(
    cum_prop = if_else(
      cum_prop > 1,
      1,
      cum_prop
    )
  )

```

## Export bead data for Days 1-3

```{r}
register <- readRDS(
  here::here("./data/colony_name_register.RDS")
) |> 
  as_tibble()

colonies_in_study <- read.csv(
  here::here(
    "./results/tables/01_table_colony_metadata.csv"
  )
) |> 
  as_tibble()

bead_data <- colonies_in_study |>
# colonies_in_study |> 
  select(
    new_name = colonyID,
    yr = which_yrs
  ) |> 
  separate_longer_delim(
    yr,
    delim = ", "
  ) |> 
  left_join(
    register,
    join_by(new_name)
  ) |> 
  rename(
    colonyID = old_name
  ) |> 
  left_join(
    tidydat |> 
      mutate(
        yr = lubridate::year(date) |> as.character()
      ),
    join_by(colonyID, yr)
  ) |> 
  mutate(
    day_num = readr::parse_number(day)
  ) |> 
  group_by(
    colonyID,
    rep
  ) |> 
  mutate(
    date = date[1] + day_num - 1,
    cum_prop = if_else(
      day_num > 1 & cum_prop < dplyr::lag(cum_prop),
      dplyr::lag(cum_prop),
      cum_prop
    )
  ) |> 
  ungroup() |> 
  select(
    - day_num
  )

# Calculate proportion removed on each day of obs ----
bead_data <- bead_data |> 
  group_by(
    colonyID,
    rep
  ) |> 
  mutate(
    prop = case_when(
      day == "D1" ~ cum_prop,
      day == "D2" ~ cum_prop - cum_prop[1],
      day == "D3" ~ cum_prop - cum_prop[2],
      .default = NA
    ) |> 
      round(2),
    .before = cum_prop
  ) |> 
  ungroup()

bead_data |> 
  mutate_all(as.factor) |> 
  summary()

bead_data |> 
  group_by(yr) |> 
  reframe(
    n_colonies = length(unique(colonyID))
  )
  
```

### Export 

```{r}
# bead_data |>
#   rowwise() |> 
#   mutate(
#     bead_color = stringr::str_split(bead_color, "-")[[1]][2]
#   ) |> 
#   ungroup() |> 
#   mutate(
#     new_name = factor(
#       new_name,
#       levels = paste0(
#         "C-",
#         bead_data$new_name |> 
#           unique() |> 
#           stringr::str_replace("C-", "") |> 
#           as.integer() |> 
#           sort()
#       )
#     ),
#     cum_prop = round(cum_prop, 3)
#   ) |> 
#   # glimpse()
#   arrange(yr, new_name, rep) |> 
#   # view()
#   write.csv(
#     "results/raw_data/bead_data_3days.csv",
#     row.names = FALSE
#   )
```

## Plot proportion of beads returned

#### Plot 1: ( day of obs X year )

```{r bead_figure, fig.width=8, fig.height=4}

# bead_data |> 
#   glimpse()
dat_beads_p1 <- bead_data |>
# bead_data |>
  group_by(
    yr,
    day
  ) |> 
  reframe(
    n = n(),
    mean = mean(cum_prop, na.rm = TRUE),
    se = sd(cum_prop, na.rm = TRUE) / sqrt(n()),
    ci = 1.96 * ( sd(cum_prop, na.rm = TRUE) / sqrt(n()) )
  ) |> 
  # glimpse()
  mutate(
    yr  = factor(yr, levels = c("2021", "2022")),
    day = factor(day, levels = paste0("D", 1:3)),
    across(
      is.numeric,
      ~ round(.x, 3)
    )
  )

dat_beads_p1 |> 
  DT::datatable()
  

beads_p1 <- dat_beads_p1 |> 
  ggplot(
    aes(
      x = day,
      y = mean
    )
  ) +
  geom_boxplot(
    width = .7,
    fatten = 1.75,
    lwd = 0.75,
    # alpha = 0.8,
    fill = "grey70",
    aes(
      # fill = day,
      x = day,
      y = cum_prop
    ),
    data = bead_data
  ) +
  # geom_errorbar(
  #   aes(
  #     ymin = mean - ci,
  #     ymax = mean + ci
  #   ),
  #   width = 0,
  #   linewidth = 2
  # ) +
  # geom_point(
  #   shape = 21,
  #   aes(
  #     fill = day
  #   ),
  #   size = 6
  # ) +
  scale_y_continuous(
    limits = c(0, 1),
    # breaks = c(0, 0.2, 0.5, 0.8, 1),
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    labels = c("0", "", "0.5", "", "1")
    # labels = c("0", "0.5", "1")
  ) +
  labs(
    y = "Proportion of beads \nremoved (cumulative)",
    x = "Day"
  ) +
  facet_grid(~ yr) +
  # facet_grid( yr ~ .) +
  theme_Publication(25) +
  theme(
    legend.position = "none"
  ) +
  scale_color_manual(
    values =  age_colors
  ) +
  scale_fill_manual(
    values =  age_colors
  )
  
```

#### Plot 2: ( day of obs X rep X yr )

```{r}

summ <- bead_data |>
# bead_data |>
  group_by(
    yr,
    colonyGroup,
    rep,
    day,
    date
  ) |> 
  reframe(
    n = n(),
    mean = mean(prop, na.rm = TRUE),
    sd = sd(prop, na.rm = TRUE),
    ci = 1.96 * ( sd(prop, na.rm = TRUE) / sqrt(n()) )
  )

# DO NOT OVERWRITE: ----
# Will remove the trial column (which was manually input)
# summ |> 
#   write.csv(
#     "results/raw_data/summ_bead_data_3days.csv",
#     row.names = FALSE
#   )
#   

summ_mod <- summ |>
  as_tibble() |> 
  mutate(
    date = lubridate::as_date(date)
  ) |> 
  mutate(
    across(
      c(mean, ci, sd),
      ~ round(.x, 3)
    )
  ) |> 
  arrange(
    yr,
    rep,
    date
  ) |> 
  left_join(
    bead_data |> 
      select(
        yr,
        colonyGroup,
        rep,
        bead_color
      ) |> 
      mutate(
        bead_color = case_when(
          stringr::str_detect(bead_color, "green") ~ "green",
          stringr::str_detect(bead_color, "pink") ~ "pink",
          stringr::str_detect(bead_color, "blue") ~ "blue",
          .default = NA
        ) |> 
          factor(
            levels = c("green", "pink", "blue")
          )
      ) |> 
      distinct(),
    join_by(yr, colonyGroup, rep)
  ) |> 
  group_by(
    yr
  ) |> 
  mutate(
    doe = (date - date[1] + 1) |> 
      as.numeric("days"),
    .after = date
  ) |> 
  ungroup()
```

## Summary Stats: Figure 1

```{r}
## [ View ] Fig 1: Supporting summary statistics ------
summ_mod |> 
  DT::datatable()

# ## [ Export ] Fig 1: Supporting summary statistics ------
# summ_mod |> 
#   rename(
#     day_of_experiment = doe,
#     n_colonies = n,
#     day_of_replicate = day,
#     replicate = rep,
#     year = yr
#   ) |> 
#   relocate(
#     day_of_experiment,
#     date,
#     .before = colonyGroup
#   ) |> 
#   glimpse()
  # write.csv(
  #   here::here(
  #     "./manuscript/resubmission/99_decision/supp_files/new_supp_summary_stats_table_for_figure1.csv"
  #   ),
  #   row.names = FALSE
  # )

```


## Kruskal Wallis: Figure 1

```{r}
fig1_data <- bead_data |>
  select(
    yr,
    colonyID,
    colonyGroup,
    rep,
    day,
    prop,
    cum_prop
  ) |> 
  glimpse()

kruskal.test(
  prop ~ day, 
  data = fig1_data |> 
    filter(
      yr == "2021" &
        rep == "rep1" &
        day != "D1"
    )
)
```


## PLOT

```{r}

# PLOT

# write a breaks func
breaks_fun <- function(x) {
  if (max(x) > 10) {
    seq(1, max(x), 1)
  } else {
    seq(1, max(x), 1)
  }
}

beads_p2a <- summ_mod |>
# summ_mod |>
  rename(
    col_set = colonyGroup,
    trial = rep
  ) |> 
  filter(
    yr == "2021"
  ) |> 
  ggplot(
    aes(
      x = doe,
      y = mean,
      group = trial
    )
  ) +
  geom_line(
    linewidth = 1.6,
    # color = "grey30"
    aes(
      col = col_set
      # lty = col_set
    )
  ) +
  geom_errorbar(
    aes(
      # ymin = mean - ci,
      ymin = mean - sd,
      # ymax = mean + ci
      ymax = mean + sd
    ),
    width = 0,
    linewidth = .5
  ) +
  geom_point(
    # shape = 21,
    stroke = 1.5,
    fill = "grey70",
    # alpha = 0.85,
    aes(
      shape = bead_color
    ),
    size = 4
  ) +
  # geom_area(
  #   aes(
  #     ymin = mean - ci,
  #     ymax = mean + ci
  #   )
  # ) +
  facet_grid(col_set ~ yr, scales = "free") +
  scale_y_continuous(
    limits = c(-0.2, 1),
    # breaks = c(0, 0.2, 0.5, 0.8, 1),
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    # labels = c("", "0.2", "0.5", "0.8", "")
    labels = c("0", "", "0.5", "", "1")
  ) +
  scale_x_continuous(
    breaks =  breaks_fun,
    limits = c(1, NA)
  ) +
  # scale_x_date(
  #   date_breaks = "4 day",
  #   date_minor_breaks = "2 day",
  #   date_labels = "%d"
  # ) +
  theme_Publication(25) +
  scale_fill_manual(
    values = age_colors,
    "Day of experiment"
  ) +
  scale_color_manual(
    values = c(
      "grey60",
      "grey60",
      # "blue",
      "black"
    ),
    name = NULL,
    labels = NULL,
    guide = NULL
  ) +
  scale_shape_manual(
    values = c(21, 23, 24),
    name = "Bead color"
  ) +
  scale_linetype_manual(
    values = c("solid", "dashed")
  ) +
  labs(
    x = "Day",
    y = "Proportion of\nbeads removed",
    title = "2021"
  ) +
  theme(
    legend.position = "none",
    # strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_blank(),
    panel.border = element_rect(colour = "black")
  )

beads_p2b <- summ_mod |>
# summ_mod |>
  rename(
    col_set = colonyGroup,
    trial = rep
  ) |> 
  filter(
    yr == "2022"
  ) |> 
  mutate(
    doe = if_else(
      trial == "rep8",
      glue::glue("{doe} ({day})") |> as.character(),
      doe |> as.character()
    )
  ) |> 
  ggplot(
    aes(
      x = as.factor(doe),
      y = mean,
      group = trial
    )
  ) +
  geom_line(
    linewidth = 1.6,
    # color = "grey30"
    aes(
      col = col_set
    )
  ) +
  geom_errorbar(
    aes(
      # ymin = mean - ci,
      ymin = mean - sd,
      # ymax = mean + ci
      ymax = mean + sd
    ),
    width = 0,
    linewidth = .5
  ) +
  geom_point(
    # shape = 21,
    stroke = 1.5,
    fill = "grey70",
    # alpha = 0.85,
    aes(
      shape = bead_color
    ),
    size = 4
  ) +
  # geom_area(
  #   aes(
  #     ymin = mean - ci,
  #     ymax = mean + ci
  #   )
  # ) +
  facet_grid(yr ~ trial, scales = "free") +
  scale_y_continuous(
    limits = c(0, 1),
    # breaks = c(0, 0.2, 0.5, 0.8, 1),
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    # labels = c("", "0.2", "0.5", "0.8", "")
    labels = c("0", "", "0.5", "", "1")
  ) +
  # scale_x_continuous(
  #   breaks =  breaks_fun,
  #   limits = c(1, NA)
  # ) +
  # scale_x_date(
  #   date_breaks = "4 day",
  #   date_minor_breaks = "2 day",
  #   date_labels = "%d"
  # ) +
  theme_Publication(25) +
  scale_fill_manual(
    values = age_colors,
    "Day of experiment"
  ) +
  scale_color_manual(
    values = c(
      "grey60",
      "grey60",
      "black"
    ),
    name = NULL,
    labels = NULL,
    guide = NULL
  ) +
  scale_shape_manual(
    values = c(21, 23, 24),
    name = "Bead color"
  ) +
  labs(
    x = "Day",
    y = "Proportion of\nbeads removed",
    title = "2022"
  ) +
  theme(
    legend.position = "none",
    # strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_blank(),
    panel.border = element_rect(colour = "black")
  )
```


```{r, fig.width=8, fig.height=14}
(
  beads_p1 / beads_p2a / beads_p2b
) + plot_annotation(
  tag_levels = 'A'
) + theme(
  plot.tag.position = c(0, 1)
)
```


# FORAGING


## Foraging data 2021

```{r get_foraging_data}
## Load the foraging data for 2021 dataset
filename3 <- paste0(
  path_to_repo,
  "/data/foraging_data.csv"
)

tidy.for.2021 <-
 read.csv(filename3) %>% 
  as_tibble() %>%
  mutate(
    date = as.Date(date, format="%m/%d/%y"),
    colonyID = as.character(colonyID),
    foraging_30s = foraging_30s / 2
  ) %>%
  # doing something, notice
  group_by(colonyID) %>% 
  mutate(
    max_foraging_30s = max(foraging_30s, na.rm = T),
    prop_foraging_30s = round(foraging_30s/max_foraging_30s, 2)
  ) |> 
  ungroup() |> 
  # take max foraging data when duplicates exist
  group_by(colonyID, date) |> 
  arrange(
    desc(prop_foraging_30s)
  ) |> 
  slice(1) |> 
  ungroup() |> 
  glimpse()
```


## Foraging data 2022

```{r foraging_2022}
# set path to your working directory (change, if necessary)
# path_to_repo2 = "/Users/biplabendudas/Documents/05_postdoc/Deborah_Gordon/04_research/2022/01_field_season_2022/04_field_work/02_rprojects/01_checking_data"
path_to_repo2 = paste0(
  path_to_repo,
  "/../01_field_season_2022/04_field_work/02_rprojects/01_checking_data"
)

# set path to the behavioral data within your repository 
path_to_data2 = "/01_data/"

## Specify the name of the master file as saved on your computer
filename2 <- "arizona_2022_master_list_Aug29_1440h.csv"

column.names <- c("date",
                  "was_it_raining",
                  "observer",
                  "colonyID",
                  "first_obs",
                  "second_obs",
                  "time",
                  "in_1",
                  "out_1",
                  "in_2",
                  "out_2",
                  "in_3",
                  "out_3",
                  "in_4",
                  "out_4",
                  "in_5",
                  "out_5",
                  "in_6",
                  "out_6",
                  "in_7",
                  "out_7",
                  "in_8",
                  "out_8"
)

tidy.for.2022 <-
  read.table(
    paste0(path_to_repo2, path_to_data2, filename2),
    sep = ",",
    header = T,
    na.strings = c(""," "),
    stringsAsFactors = F,
    skip = 1
  ) %>% 
  as_tibble() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  select(1:length(column.names)) |> 
  setNames(column.names) |> 
  # remove columns containing first and second obs
  select(-first_obs, -second_obs, -was_it_raining) |> 
  group_by(date, observer, colonyID, time) %>% 
  mutate(
    tot_in = rowSums(
      across(matches("^in_")),
      na.rm = TRUE
    ),
    tot_out = rowSums(
      across(matches("^out_")),
      na.rm = TRUE
    )
  ) %>% 
  ungroup() |>
  mutate(
    foraging_30s = (tot_in + tot_out) / 2
  ) %>% 
  select(
    date,
    colonyID,
    foraging_30s
  ) |> 
  group_by(colonyID) %>% 
  mutate(
    max_foraging_30s = max(foraging_30s, na.rm = T),
    prop_foraging_30s = round(foraging_30s/max_foraging_30s, 2)
  ) |> 
  ungroup() |> 
  # take max foraging data when duplicates exist
  group_by(colonyID, date) |> 
  arrange(
    desc(prop_foraging_30s)
  ) |> 
  slice(1) |> 
  ungroup() |> 
  glimpse()

# ## add prop_foraging data
# tidy.for.2022 <-
#   for.dat %>% 
#   select(date, 
#          colonyID,
#          foraging_30s) %>% 
#   distinct() %>% 
#   mutate(date = as.Date(date, format="%m/%d/%y")) %>%
#   mutate(colonyID = as.factor(colonyID)) %>% 
#   # doing something, notice
#   group_by(colonyID) %>% 
#   summarise(max_foraging_30s = max(foraging_30s, na.rm = T)) %>% 
#   ungroup() %>% 
#   right_join(for.dat, by="colonyID") %>% 
#   mutate(prop_foraging_30s = foraging_30s/max_foraging_30s) %>% 
#   mutate_if(is.numeric, round, 2) %>% 
#   mutate(date = as.Date(date))
# 
# 
# for.tidydat.2022 <-
#   tidydat.2022.sub %>% 
#   mutate(colonyID = as.factor(colonyID)) %>%
#   mutate(obs_day = parse_number(as.character(counted_when))) %>%
#   mutate(date = as.Date(date) + (obs_day - 1)) %>% 
#     
#   left_join(tidy.for.2022, by=c("colonyID", "date")) %>% 
#   select(colonyID:bead_color, obs_day, max_foraging_30s, foraging_30s, prop_foraging_30s) %>% 
#   
#   # mutate(colonyID = as.double(colonyID)) %>%
#   right_join(tidydat.2022.sub, by=c("colonyID", "date", "bead_color")) %>% 
#   
#   group_by(colonyID,bead_color) %>%  
#   # mutate(foraging_30s = foraging_30s[1]) %>% 
#   # mutate(prop_foraging_30s = prop_foraging_30s[1]) %>% 
#   mutate(obs_day = paste0("D",obs_day[1])) %>% 
#   ungroup() %>% 
#   
#   # mutate(s_cum_prop = round(cum_prop/foraging_30s, 2)) %>%
#   # mutate(s2_cum_prop = round(cum_prop/prop_foraging_30s, 2)) %>%
#   
#   # na.omit() %>% 
#   # filter(s_cum_prop <= 1) %>%
#   
#   select(colonyID:bead_color,
#          age_obs,age_class, 
#          beads_collected,
#          max_foraging_30s,
#          foraging_30s,
#          prop_foraging_30s,
#          counted_when, cum_prop)
```

## Combine the 2021 and 2022 foraging data

```{r}
for.tidydat <- bind_rows(tidy.for.2021, tidy.for.2022) |> 
  glimpse()
```


# Combine foraging and bead data (Day 1 only)

```{r combine_for_data}
# for.tidydat.2021 %>% summary() # 9 rows have missing foraging data | out of 46 rows
# for.tidydat.2022 %>% summary() # 12 rows have missing data | out of 156 rows

s.bead_data <-
  bead_data |> 
    filter(
      day == "D1"
    ) |> 
  left_join(
    for.tidydat,
    join_by(
      date,
      colonyID
    )
  ) |> 
  glimpse()

```

## Export combined data

```{r}
# s.bead_data |> 
#   rowwise() |>
#   mutate(
#     bead_color = stringr::str_split(bead_color, "-")[[1]][2]
#   ) |>
#   ungroup() |>
#   mutate(
#     new_name = factor(
#       new_name,
#       levels = paste0(
#         "C-",
#         bead_data$new_name |>
#           unique() |>
#           stringr::str_replace("C-", "") |>
#           as.integer() |>
#           sort()
#       )
#     ),
#     cum_prop = round(cum_prop, 3),
#     across(
#       matches("foraging"),
#       ~ if_else(
#         is.na(.x),
#         "missing",
#         as.character(.x)
#       )
#     )
#   ) |>
#   arrange(yr, new_name, rep) |> 
#   # view()
#   write.csv(
#     "results/raw_data/foraging_bead_data_day1.csv",
#     row.names = FALSE
#   )
```


# SUMMARY OF COLONIES USED (BEADS)

```{r}
bead_data |> 
  group_by(
    rep,
    colonyID
  ) |> 
  slice(1) |> 
  ungroup() |> 
  mutate(
    across(
      c(colonyID, new_name),
      ~ as.character(.x) |> 
        stringr::str_replace(
          "C-",
          ""
        ) |> 
        as.integer()
    )
  ) |> 
  glimpse() |> 
  group_by(bead_color, colonyGroup, date) |> 
  reframe(
    n = n(),
    colonyID = paste(sort(new_name), collapse = ", ")
    # old_names = paste(sort(colonyID), collapse = ", ")
  ) |> 
  # view()
  # write.csv(
  #   here::here(
  #     "./manuscript/resubmission/00_figures_tables/01_supp_colonies_used_summary.csv"
  #   ),
  #   row.names = FALSE
  # )
  DT::datatable()

```

## Save compiled data

```{r save_compiled_data_v1}
saveRDS(
  s.bead_data, 
  file = paste0(
    path_to_repo,
    "/data/compiled_data_s_bead_data_mod_2021_2022"
  )
)
```


### What does each column mean?

The columns in the above dataset has the following meaning.

- *yr*: which year?
    - year in which the replicate was performed; n=2; 2021 and 2022
- *date*: date of experiment
    - either the first date of the experimental replicate is shown
    - or exact dates are shown
- *rep*: which replicate?
    - unique replicates were performed; n=6; rep1 through rep6
    - color of the bead that was used; n=3; green/pink/blue
    - year in which the replicate was performed; n=2; 2021(21) and 2022

- *colonyID*: which colony?
    - unique identifiers (name) of the colony used
    - n = 36 unique colonies across the two years
    - 13 in 2021 and 19 in 2022; 6 colonies in both

- *beads_collected*: number of beads collected
    - usually 50 are oil-coated beads are presented
- *counted_when*: which day-time for counting number of returned beads
    - Day 1 or 2 (D1 or D2)
    - before noon (bN), around 11 am
    - Morning (M), around 8 AM
- *cum_prop*: proportion of beads returned on Day 1 or 2 (D1 or D2)
    - proportion was calculated as (# beads returned)/(# beads collected)
    - Day 1 observation of beads returned was performed before noon (bN), around 11 am
    - Day 2 observations were performed either in the morning (M), around 8 AM, or before noon

- *age_obs*: age of the colony at observation
    - units: years
- *age_class*: age-class of the colony
    - age class that a colony belongs to (n=3)
    - classified as 1-2yr v. 3-4yr v. 5+ year old

- *max_foraging_30s*: maximum foraging rate observed for a colony
    - units: ants per 30 seconds
    - maximum rate observed for the colony across all observation days in a year
    - seasonal, in our case yearly, measure
- *foraging_30s*: foraging rate observed for a colony on the given day
    - units: ants per 30 seconds
    - rate observed during the morning activity period on the day of the experiment
    - daily measure


# Year to Year Difference


### Variation across years (D1 only)

```{r fig_v4, fig.width=9, fig.height=5}
# fig10a <-
#   dat.d1 %>%
#   ggplot(aes(x=yr, y=cum_prop1)) +
#   geom_boxplot(size=1, alpha=0.5, aes(fill=yr)) +
#   geom_jitter(size=4, alpha=0.35, width = 0.1) +
#   theme_Publication(25) +
#   theme(axis.title.x = element_blank(),
#         legend.position = "none") +
#   scale_y_continuous(n.breaks = 3) +
#   scale_fill_manual(values = yr_colors) +
#   labs(x="year",
#        y="proportion of beads",
#        title = "Day 1")

foo <- s.bead_data |> 
  filter(
    !is.na(foraging_30s)
  )

foo |> 
  select(
    yr,
    cum_prop,
    foraging_30s
  ) |> 
  tidyr::pivot_longer(
    !yr,
    names_to = "what",
    values_to = "value"
  ) |> 
  mutate(
    what = if_else(
      what == "cum_prop",
      "Proportion of beads",
      "Foraging activity"
    ) |> 
      factor(
        levels = c(
          "Proportion of beads",
          "Foraging activity"
        )
      )
  ) |> 
  ggplot(
    aes(
      x = yr,
      y = value
    )
  ) +
  geom_boxplot(
    size = 1, alpha = 0.85, fill = "grey", color = "black",
    outlier.color = NA
  ) +
  geom_jitter(
    size = 4, alpha = 0.35, width = 0.15
  ) +
  facet_wrap(~ what, scales = "free_y") +
  theme_Publication(25) +
  theme(
    axis.title.x = element_blank(),
    title = element_blank(),
    legend.position = "none"
  ) +
  scale_y_continuous(n.breaks = 4) +
  labs(
    y = ""
  )

```


# Variation in abiotic conditions

```{r abiotic_plot_v1, fig.width=9, fig.height=4}
clim.dat <- read.csv(
  paste0(
    path_to_repo,
    "/data/all_climate_data_beadexp_2021_2022.csv"
  )
)

writeLines("How many unique dates, in each year, are we 
           using to calculate mean and 95% CI?")

clim.dat %>%
  select(1) %>%
  distinct() %>%
  mutate(date = as.character(date)) %>%
  group_by(date) %>%
  mutate(yr = str_split_fixed(date, n=2, pattern = "-")[[1]][1]) %>%
  mutate(day = str_split_fixed(date, n=3, pattern = "-")[[3]][1]) %>%
  ungroup() %>%
  group_by(yr) %>%
  summarise(n.days = length(unique(day)),
            which.month = "August",
            which.days = paste(unique(day), collapse = ", "))

clim.dat_summ <-
  clim.dat %>%
    as_tibble() %>%
  select(date, time, temp_F, rH, precip=precip_accum_inch) %>%
  distinct() %>%
  mutate(temp_C=(temp_F-32)*(5/9)) %>%
  group_by(date) %>%
  mutate(yr = str_split(date, "-")[[1]][1]) %>%
  ungroup() %>%
  mutate(yr = as.factor(yr)) %>%
  pivot_longer(cols = temp_F:temp_C,
               names_to = "what",
               values_to = "value") %>%

  # ggplot(aes(x=what, y=value, fill=yr)) +
  # scale_color_manual(values = yr_colors) +
  # scale_fill_manual(values = yr_colors) +
  # geom_boxplot(width=.4, position = position_dodge(width = .60)) +
  # theme_Publication(20) +
  # theme(legend.position = "none",
  #       axis.title.x = element_blank(),
  #       axis.title.y = element_blank())

  filter(what != "temp_F") %>%

  group_by(yr, what) %>%
  summarize(mean.val = round(mean(value, na.rm = T),2),
            sd.val = round(sd(value, na.rm = T),2),
            n.val = n()) %>%
  mutate(se.val = round(sd.val/sqrt(n.val), 2),
         ci95 = round(qt(1 - (0.05/2), n.val - 1)*se.val, 2)) %>%
  ungroup()

clim.dat_summ |> 
  arrange(what, yr) |> 
  DT::datatable()

clim.dat_summ %>%
    ggplot(aes(x=yr, y=mean.val, col=yr)) +
    geom_errorbar(
      aes(ymin=mean.val-ci95, ymax=mean.val+ci95), 
      width=0.4, 
      size=2, 
      col="black"
    ) +
    geom_point(size=4, position = position_dodge(width = 2), col = "red") +
    facet_wrap(~as.factor(what), scales = "free_y") +
    theme_Publication(base_size = 22) +
    scale_color_manual(values = yr_colors) +
    scale_fill_manual(values = yr_colors) +
    # scale_shape_manual(values = c(17,17,16, 16,16,16)) +
    # coord_flip() +
    theme(legend.position = "none",
          axis.title.x = element_blank()) +
    labs(x="",
         y="",
         subtitle = "precip inch; humidity %rH; temperature ºC")
# )

```

## Kruskal Wallis Tests

```{r}
purrr::map(
  c("temp_F", "rH", "precip_accum_inch"),
  function(x) {
    foo <- clim.dat |>
      mutate(
        yr = lubridate::ymd(date) |> lubridate::year()
      )
    kruskal.test(foo[[x]] ~ foo$yr) |> print()
  }
) |> 
  setNames(
    c("temp_F", "rH", "precip_accum_inch")
  )

```


# AGE

```{r, fig.width=9, fig.height=5}
bead_data |> 
  filter(
    yr == "2022"
  ) |> 
  ggplot(
    aes(x = age_class, y = cum_prop, fill = age_class)
  ) +
  facet_grid(~ day) +
  geom_boxplot() +
  geom_jitter(
    size = 3,
    width = 0.1,
    alpha = 0.3
  ) +
  scale_y_continuous(
    n.breaks = 3
  ) +
  labs(
    x = "Age class",
    y = "Proportion of beads\n(cumulative)"
  ) +
  scale_fill_manual(values = viridis::viridis(5)[c(2,4,5)]) +
  theme_Publication(20) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom"
  )
  

```

## Kruskal Wallis Test 

Comparing the medians across different age groups

```{r}
bead_data |> 
  filter(yr == "2022") |> 
  group_by(day) |> 
  reframe(
    statistic = kruskal.test(cum_prop ~ age_class)$statistic,
    p_value = kruskal.test(cum_prop ~ age_class)$p.value,
    df = kruskal.test(cum_prop ~ age_class)$parameter
  ) |> 
  mutate_if(
    is.numeric,
    ~ round(.x, 3)
  ) |> 
  print()
```


# 1. Rate of rubbish removal

```{r}
bead_data |> 
  group_by(yr, day) |> 
  reframe(
    n = n(),
    mean_value = mean(cum_prop, na.rm = TRUE),
    median_value = median(cum_prop, na.rm = TRUE),
    se = sd(cum_prop, na.rm = TRUE) / sqrt(n),
    lower_ci = t.test(cum_prop)$conf.int[1],
    upper_ci = t.test(cum_prop)$conf.int[2]
  ) |> 
  kableExtra::kable()
```

# 3. Relation between bead returns and foraging

## Prep data

```{r}
dat.mod <- foo |>
  mutate(
    beads_returned = round(
      prop*beads_collected, 
      0
    )
  ) |> 
  group_by(
    colonyID,
    yr
  ) |> 
  mutate(
    n_reps_yr = n()
  ) |> 
  ungroup() |> 
  filter(
    n_reps_yr > 1
  )

```


## Evaluate models

```{r, include=FALSE}

# GLMM --------------------------------------------------------------------
mm0 <- lme4::glmer(
  cbind(
    beads_returned,
    beads_collected - beads_returned
  ) ~ (1|colonyID/rep),
  # data = dat.mod, 
  data = dat.mod |> 
    mutate(
      yr = factor(
        yr,
        levels = c("2021", "2022")
      )
    ),
  family = binomial(link = "logit")
)

mm1 <- lme4::glmer(
  cbind(
    beads_returned,
    beads_collected - beads_returned
  ) ~ (1|colonyID/rep) +
    yr +
    prop_foraging_30s +
    yr:prop_foraging_30s,
  # data = dat.mod, 
  data = dat.mod |> 
    mutate(
      yr = factor(
        yr,
        levels = c("2021", "2022")
      )
    ),
  family = binomial(link = "logit")
)

anova(mm1, mm0)
```

## Fixed effects

```{r}
# Re-check importance of random effects
writeLines("Proposed MODEL (m1):")
summary(mm1)[["call"]] |> 
  print()

writeLines("Interaction term")
mm2 <- update(
  mm1,
  . ~ . - yr:prop_foraging_30s
)
anova(mm2, mm1)

writeLines("Normalized foraging")
mm3 <- update(
  mm2,
  . ~ . - prop_foraging_30s
)
anova(mm3, mm2)

```


## Random effects

```{r}
writeLines("Random: Colony ID")
mm3a <- update(
  mm3,
  . ~ . - (1|colonyID/rep) + (1|rep)
)
anova(mm3a, mm3)

writeLines("Random: Replicate")
mm3b <- update(
  mm3,
  . ~ . - (1|colonyID/rep) + (1|colonyID)
)
anova(mm3b, mm3)

```


## Explore best model

```{r, fig.width=6, fig.height=5}
## Specify model ----
glmm <- mm3

## QC model ----
plot(glmm)

## Export summary table ----
summary(glmm) |> 
  print()

# summary(glmm) |> 
#   purrr::pluck(coefficients) |> 
#   as.data.frame() |> 
#   tibble::rownames_to_column(
#     var = "Variable"
#   ) |> 
#   as_tibble() |> 
#   janitor::clean_names() |> 
#   mutate_if(is.numeric, ~ round(.x, 4)) |> 
#   DT::datatable()
  # write.csv(
  #   "./results/tables/02_resubmission_model_summary.csv",
  #   row.names = FALSE
  # )

```

## Backtransformed from logit scale 

Summary Table with odds-ratio instead of log-odds

```{r}
# Tidy the model output
tidy_summary <- broom.mixed::tidy(
# broom.mixed::tidy(
  glmm,
  effects = "fixed",
  conf.int = TRUE,
  exponentiate = TRUE
) |> 
  mutate(
    across(
      is.numeric,
    ~ round(.x, 3)
    )
  ) |> 
  select(
    term,
    estimate,
    matches("^conf."),
    statistic,
    p.value
  )

# # Print the results
# print(tidy_summary)

# Print the summary table
tidy_summary |> 
  mutate(
    across(
      estimate:statistic,
      ~ round(.x, 2)
    )
  ) |> 
  DT::datatable()
```


# Conclusion

Overall, these findings highlight the importance of foraging behavior in the likelihood of waste removal, and suggests that year to year differences in abiotic conditions (effects of year) can modulate the relation between normalized foraging rate (proportion of maximum possible) and the likelihood of waste removal in harvester ant colonies.

> RANDOM EFFECTS: The likelihood of bead returns also varies significantly across days of experiment (replicate) and across colonies (colonyID).


<!-- ```{r, fig.width=7, fig.height=4} -->

<!-- ## plot predictions ---- -->
<!-- ### bead returns ~ prop foraging ---- -->
<!-- range_foraging <- dat.mod |>  -->
<!--   pull(prop_foraging_30s) |>  -->
<!--   # pull(foraging_30s) |>  -->
<!--   range() -->

<!-- # seq_foraging <- seq( -->
<!-- #   # range_foraging[1], range_foraging[2], 10 -->
<!-- #   range_foraging[1], range_foraging[2], 10 -->
<!-- # ) -->

<!-- seq_foraging <- c(0, 0.25, 0.5, 0.75, 1) -->

<!-- pred.glmm <- emmeans::emmeans( -->
<!--   glmm, -->
<!--   c("yr", "prop_foraging_30s"), -->
<!--   # c("yr", "foraging_30s"), -->
<!--   at = list( -->
<!--     # foraging_30s = seq_foraging -->
<!--     prop_foraging_30s = seq_foraging -->
<!--   ), -->
<!--   # at = list(prop_foraging_30s = c(0, 0.25, 0.5, .75, 1)), -->
<!--   type = "response" -->
<!-- ) |>  -->
<!--   broom.mixed::tidy() |>  -->
<!--   janitor::clean_names() |>  -->
<!--   # filter( -->
<!--   #  p_value < 0.05  -->
<!--   # ) |>  -->
<!--   rename( -->
<!--     prop_beads_returned = prob -->
<!--   ) -->

<!-- p1 <- foo |>  -->
<!--   rename( -->
<!--     prop_beads_returned = prop -->
<!--   ) |>  -->
<!--   ggplot( -->
<!--     aes( -->
<!--       # x = foraging_30s, -->
<!--       x = prop_foraging_30s, -->
<!--       # x = log2(age_obs), -->
<!--       # x = yr, -->
<!--       y = prop_beads_returned -->
<!--       # y = qlogis(prop_beads_returned) # logit scale -->
<!--     ) -->
<!--   ) + -->
<!--   geom_point( -->
<!--     shape = 21, -->
<!--     color = "black", -->
<!--     fill = "black", -->
<!--     # aes( -->
<!--     #   fill = yr -->
<!--     # ), -->
<!--     size = 3, -->
<!--     alpha = 0.7 -->
<!--   ) + -->
<!--   # scale_x_continuous( -->
<!--   #   breaks = c(0, .5, 1), -->
<!--   #   labels = c(0, 0.5, 1) -->
<!--   # ) + -->
<!--   # scale_y_continuous( -->
<!--   #   breaks = c(0, .5, 1), -->
<!--   #   labels = c(0, 0.5, 1) -->
<!--   # ) + -->
<!--   # scale_color_manual( -->
<!--   #   values = c("grey20", "orange") -->
<!--   # ) + -->
<!--   # scale_fill_manual( -->
<!--   #   values = c("grey20", "orange") -->
<!--   # ) + -->
<!--   theme_bw(24) + -->
<!--   theme( -->
<!--     legend.position = "bottom" -->
<!--   ) -->

<!-- p2a <- p1 + -->
<!--   geom_ribbon( -->
<!--     aes( -->
<!--       ymin = prop_beads_returned - std_error*qnorm(0.975), -->
<!--       ymax = prop_beads_returned + std_error*qnorm(0.975) -->
<!--     ), -->
<!--     fill = "grey", -->
<!--     alpha = 0.35, -->
<!--     data = pred.glmm -->
<!--   ) + -->
<!--   geom_line( -->
<!--     data = pred.glmm, -->
<!--     col = "black", -->
<!--     # aes(col = yr), -->
<!--     size = 2 -->
<!--   ) + -->
<!--   labs( -->
<!--     y = "Proportion of\nbeads removed", -->
<!--     x = "Normalized foraging rate" -->
<!--   ) + -->
<!--   scale_x_continuous( -->
<!--     breaks = c(0, .50, 1) -->
<!--   ) + -->
<!--   scale_y_continuous( -->
<!--     breaks = c(0, .50, 1) -->
<!--   ) + -->
<!--   # labs( -->
<!--   #   title = "Predicted relation", -->
<!--   #   subtitle = "(95% Confidence Interval)" -->
<!--   # ) -->
<!--   # labs( -->
<!--   #   y = "Proportion beads returned", -->
<!--   #   x = "Proportion foraging" -->
<!--   # ) + -->
<!--   facet_grid( ~ yr) -->

<!-- print(p2a) -->
<!-- # #### Save plot ---- -->
<!-- # bd.saveplot( -->
<!-- #   name = "pred_glmm_foraging", -->
<!-- #   width = 20, -->
<!-- #   height = 14, -->
<!-- #   # the plot -->
<!-- #   plot = p2a -->
<!-- # ) -->
<!-- ``` -->

